Practical Optional Types for Clojure

### 2015-02-26 overview.tex
use formalism in judgements

This function takes a value that is either \emph{\#f} % mintinline really hates #
or a number, represented by an untagged \emph{union type}.
The `then' branch has an implicit invariant
that \rkt{x} is a number, which is automatically inferred with occurrence typing
and type checked without further annotations.
Here is same program in Typed Clojure.

@@

Symbols support \emph{metadata}, an arbitrary map of
data. The syntax \clj{^File x} is a single expression that is the symbol \clj{x} 
with metadata \clj{{:tag File}}.
In binding positions like \clj{(fn [^File x] ...)}
the metadata is preserved in occurrences of \clj{x}.

@@

\emph{Destructuring} is binding position syntax for pattern matching.
A \emph{map destructuring} contains pairs of symbols and keywords that bind the symbols
to lookups on that keyword with the current argument, optionally terminated by
\clj{:as} and a symbol, which binds the symbol to the current argument. For example
$$
\clj{(fn [{^File x :x, ^String y :y, :as m}] ...)}
$$ 
expands to
\begin{minted}{clojure}
  (fn [m___0]
    (let [m m___0,
          ^File x (:x m),
          ^String y (:y m)]
      ...))
\end{minted}
Destructuring is supported similarly by \clj{let}.

@@

\subsection{Java interoperability}

>>>>

This function takes a value that is either \emph{\#f} % mintinline really hates #
or a number, represented by an untagged \emph{union type}.
The `then' branch has an implicit invariant
that \rkt{x} is a number, which is automatically inferred with occurrence typing
and type checked without further annotations.

We chose to build on the ideas and implementation
of Typed Racket to implement a type system targeting Clojure for several reasons.
Initially, the similarities between Racket and Clojure drew us to
investigate the effectiveness of repurposing occurrence typing
for a Clojure type system---both languages share a Lisp heritage,
similar standard functions 
(for instance \clj{map}
in both languages is variable-arity)
and idioms.
While Typed Racket is gradually typed and has sophisticated
dynamic semantics for cross-language interaction, we 
chose to first implement
the static semantics
with the hope to extend Typed Clojure to be gradually typed at a future date.
Finally,
Typed Racket's combination of bidirectional checking
and occurrence typing presents a successful model for 
type checking dynamically typed programs without compromising
soundness, where other approaches like success typing~\cite{todo-dialyzer-paper}
or soft typing~\cite{successtyping-todo}
fail to prove strong properties about programs
and have proved too complicated in practice respectively.

Here is above program in Typed Clojure.

@@

Symbols support \emph{metadata}, an arbitrary map of
data. The syntax \clj{^File x}, shorthand for \clj{^{:tag File} x}, is a single expression that is the symbol \clj{x} 
with metadata \clj{{:tag File}}.
In binding positions like \clj{(fn [^File x] ...)}
the compile-time metadata is preserved in occurrences of \clj{x}
and utilized by the compiler.

@@

\emph{Destructuring} is binding position syntax for pattern matching.
A \emph{map destructuring} contains pairs of symbols and keywords that bind the symbols
to lookups on that keyword with the current argument, optionally terminated by
\clj{:as} and a symbol, which binds the symbol to the current argument. For example
$$
\clj{(fn [{^File x :x, :as m}] ...)}
$$ 
expands to a \clj{let} that binds \clj{m} to the argument and 
\clj{^File x} to \clj{(:x m)}.
Destructuring is supported similarly by \clj{let}.

@@

\subsection{Java interoperability}
\label{sec:overviewjavainterop}
